---
title: "p8105_hw2_SC6036"
output: github_document
date: "2026-10-05"
---

```{r}
library(tidyverse)
```

# Question 1
```{r}
nycTransit = 
  read_csv("data/NYCTransit.csv") |> 
  select("Line", "Station Name", "Station Latitude", "Station Longitude", 
         "Route1", "Route2", "Route3", "Route4", "Route5", "Route6", "Route7", 
         "Route8", "Route9", "Route10", "Route11", "Entry", "Vending", "Entrance Type", "ADA") |> 
  janitor::clean_names() |> 
  mutate(
    entry = case_match(
      entry, 
      "YES" ~ 1,
      "NO" ~ 2
    )
  )
nycTransit
```
This data set has 18 columns and 1,868 rows, with selected variables "Line", "Station Name", "Station Latitude", "Station Longitude", Route 1-10 "Entry", "Vending", "Entrance Type", "ADA". So far I have cleaned the variable names, selected the variables we want to look at, and used the mutate function to code 1 to "YES" and 2 to "NO". I wouldn't consider this data tidy because there are many binary variables aside from "Entry"; such as "ADA" and "Vending". There is also many NA values and I would want to look further into that to see if they can be dropped from the data set. I also would make the routes less redundant using pivot_longer(). 

```{r}
nycTransit |> 
  distinct(line, station_name) # or

distinct(nycTransit, line, station_name, keep_all = T)

filter(nycTransit, ada == "FALSE")
filter(nycTransit, ada == "TRUE")

nrow(filter(nycTransit, vending == "NO", entry == 1)) / 
  nrow(filter(nycTransit, vending == "NO"))
# no vending and entry allowed / all entrances with no vending 
```
There are 465 distinct stations. Of these, only 468 stations are ADA compliant. About 37.7% of station entrances and exists without vending allow entrance.

```{r}
nycTransitTidy <- nycTransit |>
  mutate(across(route1:route11, as.character)) |>
  pivot_longer(
    cols = starts_with("route"),
    names_to = "route_number",
    names_prefix = "route",
    values_to = "route_name"
  )
nycTransitTidy

nycTransitTidy |> 
  filter(route_name == "A") |> 
  distinct(station_name, line)

nycTransitTidy |> 
  filter(route_name == "A", ada == "TRUE") |> 
  distinct(station_name, line)
```
There are 60 distinct stations that serve the A train. Of those, 17 are ADA compliant
