---
title: "p8105_hw2_SC6036"
output: github_document
date: "2026-10-05"
---

```{r}
library(tidyverse)
library(readxl)
```

# Question 1
```{r}
nycTransit = 
  read_csv("data/NYCTransit.csv") |> 
  select("Line", "Station Name", "Station Latitude", "Station Longitude", 
         "Route1", "Route2", "Route3", "Route4", "Route5", "Route6", "Route7", 
         "Route8", "Route9", "Route10", "Route11", "Entry", "Vending", "Entrance Type", "ADA") |> 
  janitor::clean_names() |> 
  mutate(
    entry = case_match(
      entry, 
      "YES" ~ 1,
      "NO" ~ 2
    )
  )
nycTransit
```
This data set has 18 columns and 1,868 rows, with selected variables "Line", "Station Name", "Station Latitude", "Station Longitude", Route 1-10 "Entry", "Vending", "Entrance Type", "ADA". So far I have cleaned the variable names, selected the variables we want to look at, and used the mutate function to code 1 to "YES" and 2 to "NO". I wouldn't consider this data tidy because there are many binary variables aside from "Entry"; such as "ADA" and "Vending". There is also many NA values and I would want to look further into that to see if they can be dropped from the data set. I also would make the routes less redundant using pivot_longer(). 

```{r}
# distinct station
nycTransit |> 
  distinct(line, station_name) # or

distinct(nycTransit, line, station_name, keep_all = T)

# distinct ADA compliant stations 
filter(nycTransit, ada == "FALSE")
filter(nycTransit, ada == "TRUE")

# no vending and entry allowed / all entrances with no vending 
nrow(filter(nycTransit, vending == "NO", entry == 1)) / 
  nrow(filter(nycTransit, vending == "NO"))
```
There are 465 distinct stations. Of these, only 468 stations are ADA compliant. About 37.7% of station entrances and exists without vending allow entrance.

```{r}
nycTransitTidy <- nycTransit |>
  mutate(across(route1:route11, as.character)) |>
  pivot_longer(
    cols = starts_with("route"),
    names_to = "route_number",
    names_prefix = "route",
    values_to = "route_name"
  )
nycTransitTidy

# distinct stations serving A
nycTransitTidy |> 
  filter(route_name == "A") |> 
  distinct(station_name, line)

# distinct stations serving A and ADA compliant
nycTransitTidy |> 
  filter(route_name == "A", ada == "TRUE") |> 
  distinct(station_name, line)
```
There are 60 distinct stations that serve the A train. Of those, 17 are ADA compliant

# Question 2
```{r}
# Cleaning data sets 

MrTrashWheel <- read_excel("data/202509 Trash Wheel Collection Data.xlsx", sheet = "Mr. Trash Wheel", range = "A2:N709") |> 
  janitor::clean_names() |> 
  drop_na(dumpster) |> 
  mutate(
    sports_balls = as.integer(round(sports_balls)),
    trash_wheel = "Mr. Trash Wheel",
    year = as.integer(year)
  )

ProfTrashWheel <- read_excel("data/202509 Trash Wheel Collection Data.xlsx", sheet = "Professor Trash Wheel", range = "A2:M134") |> 
  janitor::clean_names() |> 
  drop_na(dumpster) |> 
  mutate(
    trash_wheel = "Professor Trash Wheel",
    year = as.integer(year)
  )

Gwynnda <- read_excel("data/202509 Trash Wheel Collection Data.xlsx", sheet = "Gwynns Falls Trash Wheel", range = "A2:L351") |> 
  janitor::clean_names() |> 
  drop_na(dumpster) |> 
  mutate(
    trash_wheel = "Gwynnda",
    year = as.integer(year)
  )
# making the year type consistent
```

```{r}
# Combining
trashWheels <- bind_rows(MrTrashWheel, ProfTrashWheel, Gwynnda) |> 
  relocate(trash_wheel)
```
The combined data set is 15 columns by 1,188 rows. Some key variables are "trash_wheel", "year", "dumpster", "homes_powered", and "cigarette_butts". 

```{r}
trashWheels |> 
  filter(trash_wheel == "Mr. Trash Wheel") |> 
  summarize(total_weight = sum(weight_tons, na.rm = T))

trashWheels |> 
  filter(trash_wheel == "Gwynnda") |> 
  filter(month == "June" & year == 2022) |> 
  summarize(cigarette_butts = sum(cigarette_butts, na.rm = T))
```
The total weight of trash collected by Mr. Trash Wheel is 2283.84 (lbs?) and in June 2022, Gwynnda colleged 18120 cigarette butts. 

## Question 3















