---
title: "p8105_hw2_SC6036"
output: github_document
date: "2026-10-05"
---

```{r}
library(tidyverse)
library(readxl)
```

# Question 1
```{r}
nycTransit = 
  read_csv("data/NYCTransit.csv") |> 
  select("Line", "Station Name", "Station Latitude", "Station Longitude", 
         "Route1", "Route2", "Route3", "Route4", "Route5", "Route6", "Route7", 
         "Route8", "Route9", "Route10", "Route11", "Entry", "Vending", "Entrance Type", "ADA") |> 
  janitor::clean_names() |> 
  mutate(
    entry = case_match(
      entry, 
      "YES" ~ 1,
      "NO" ~ 2
    )
  )
nycTransit
```
This data set has 18 columns and 1,868 rows, with selected variables "Line", "Station Name", "Station Latitude", "Station Longitude", Route 1-10 "Entry", "Vending", "Entrance Type", "ADA". So far I have cleaned the variable names, selected the variables we want to look at, and used the mutate function to code 1 to "YES" and 2 to "NO". I wouldn't consider this data tidy because there are many binary variables aside from "Entry"; such as "ADA" and "Vending". There is also many NA values and I would want to look further into that to see if they can be dropped from the data set. I also would make the routes less redundant using pivot_longer(). 

```{r}
# distinct station
nycTransit |> 
  distinct(line, station_name) # or

distinct(nycTransit, line, station_name, keep_all = T)

# distinct ADA compliant stations 
filter(nycTransit, ada == "FALSE")
filter(nycTransit, ada == "TRUE")

# no vending and entry allowed / all entrances with no vending 
nrow(filter(nycTransit, vending == "NO", entry == 1)) / 
  nrow(filter(nycTransit, vending == "NO"))
```
There are 465 distinct stations. Of these, only 468 stations are ADA compliant. About 37.7% of station entrances and exists without vending allow entrance.

```{r}
nycTransitTidy <- nycTransit |>
  mutate(across(route1:route11, as.character)) |>
  pivot_longer(
    cols = starts_with("route"),
    names_to = "route_number",
    names_prefix = "route",
    values_to = "route_name"
  )
nycTransitTidy

# distinct stations serving A
nycTransitTidy |> 
  filter(route_name == "A") |> 
  distinct(station_name, line)

# distinct stations serving A and ADA compliant
nycTransitTidy |> 
  filter(route_name == "A", ada == "TRUE") |> 
  distinct(station_name, line)
```
There are 60 distinct stations that serve the A train. Of those, 17 are ADA compliant

# Question 2
```{r}
# Cleaning data sets 

MrTrashWheel <- read_excel("data/202610 Trash Wheel Collection Data.xlsx", sheet = "Mr. Trash Wheel", range = "A2:N742") |> 
  janitor::clean_names() |> 
  drop_na(dumpster) |> 
  mutate(
    sports_balls = as.integer(round(sports_balls)),
    trash_wheel = "Mr. Trash Wheel",
    year = as.integer(year) #read in as text for consistency in the combining step
  )

ProfTrashWheel <- read_excel("data/202610 Trash Wheel Collection Data.xlsx", sheet = "Professor Trash Wheel", range = "A2:M139") |> 
  janitor::clean_names() |> 
  drop_na(dumpster) |> 
  mutate(
    trash_wheel = "Professor Trash Wheel",
    plastic_bottles = as.numeric(plastic_bottles), # one cell says "dive"
    year = as.integer(year) 
  )

Gwynnda <- read_excel("data/202610 Trash Wheel Collection Data.xlsx", sheet = "Gwynnda the Good Wheel of the W", range = "A2:L394") |> 
  janitor::clean_names() |> 
  drop_na(dumpster) |> 
  mutate(
    trash_wheel = "Gwynnda",
    year = as.integer(year)
  )
# making the year type consistent
```

```{r}
# Combining
trashWheels <- bind_rows(MrTrashWheel, ProfTrashWheel, Gwynnda) |> 
  relocate(trash_wheel)
```
The combined data set is 15 columns by 1,269 rows. Some key variables are "trash_wheel", "year", "dumpster", "homes_powered", and "cigarette_butts" with one row per dumpster brought in by each of the trash wheels. 

```{r}
trashWheels |> 
  filter(trash_wheel == "Professor Trash Wheel") |> 
  summarize(total_weight = sum(weight_tons, na.rm = T))

trashWheels |> 
  filter(trash_wheel == "Gwynnda") |> 
  filter(month == "June" & year == 2022) |> 
  summarize(cigarette_butts = sum(cigarette_butts, na.rm = T))
```
The total weight of trash collected by Professor Trash Wheel is 294.96 tons and in June 2022, Gwynnda collected 18120 cigarette butts. 

## Question 3
```{r}
MCI_baseline_df <- read_csv("data/MCI_baseline.csv", skip = 1, na = c("", "NA", "Na", ".")) |>  
    #removing all types of NA and skipping the data dictionary line
  janitor::clean_names() |> #cleaning up names
  mutate(
    sex = factor(case_match(sex, 0 ~ "Female", 1 ~ "Male")), 
      # matching 0 as female and 1 as male
    apoe4 = factor(case_match(apoe4, 0 ~ "Non-carrier", 1 ~ "Carrier")) 
      # matching 0 as non carrier and 1 as carrier
  )

## removing those who already have MCI (failing to meet criteria at baseline)
mci_baseline <- MCI_baseline_df |> 
  filter(is.na(age_at_onset) | age_at_onset > current_age) 
    # onset age at or below baseline age fails the inclusion criteria

mean(mci_baseline$current_age)

mci_baseline |>
  filter(sex == "Female") |>
  summarize(mean(apoe4 == "Carrier"))
```
There are a total of 483 participants, 479 of which are MCI free at baseline. The average age is ~65 years old. Proportion of women in the study who are APOE4 carriers is 0.3, or 30%. (63/210 women)

```{r}
mci_amyloid <- read_csv("data/mci_amyloid.csv", skip = 1, na = c("", "NA", "Na")) |> 
    # removing all types of NA, "Na" appears in a few rows
    # skip data dictionary row
  janitor::clean_names() |> #cleaning up names
  rename(id = study_id, time_0 = baseline) |> 
    # using time_0 as baseline/reference and matching "id" to join cleanly 
  pivot_longer(
    cols = time_0:time_8,
      # data is wide, making it longer
    names_to = "visit_time",
    names_prefix = "time_",
    values_to = "amyloid_ratio"
  ) |>
  mutate(visit_time = as.numeric(visit_time)) # making sure visit_time is a number rather than character
```
The data set has 487 participants or "Study ID" and 6 columns. After cleaning the dataset is 3 columns by 2435 rows. The "visit_time" column now indicates a number for the number of years. Each row is a participant per visit. 

```{r}
anti_join(mci_baseline, mci_amyloid, by = "id") |> distinct(id)
anti_join(mci_amyloid, mci_baseline, by = "id") |> distinct(id) 
```
In baseline data, there are 8 participants or ids' that have only demographics. In amlyloid data, there are 16 participants or ids' that have only biomarker data. 

```{r}
mci_combined <- inner_join(mci_baseline, mci_amyloid, by = "id")

write_csv(mci_combined, "data/mci_combined.csv")
```
After keeping only the participants in both datasets who are eligible, we are left with 471 participants. The dataset is long, with 2,355 rows and 8 columns. There are 8 variables and NA represents missing visits, although we can drop those if we want. 



